---
title: "Fine root biomass - Miscanthus x giganteus - IA"
author: "Ryan Quinn"
date: "March 7, 2020"
output: html_document
---

**Part 1: Data Cleaning and Early Exploration**

**Notes:** 
  -Earlier data exploration examined the effects of outlier removal on the model. Removing outliers removes most 0-10cm depth bm datapoints. So, we do not remove outliers for this analysis. 
  -This data IS NOT PUBLICALLY AVAILABLE and will not be until the publication is in press. However, these scripts may help someone currently processing soil C flux data. 


-load libraries
-Define functions needed throughout script
-Load data, remove missing datapoints
-Basic exploration of data distribution, shape, etc. 
```{r}
library(readr)
library(lmerTest)
library(emmeans)
library(ggplot2)
#library(viridis) #went with colorcoding by x variable instead of colorblind friendly for presentation b/c nobody on committee colorblind
library(dplyr) #summary statistics 

se.error <- function(x)sd(x)/sqrt(length(x)) #calculate SE during data aggregation 
pd <- position_dodge(width = 1) #position error bars on  figures

ia <- read_csv("C:/Users/ryan_/OneDrive/Desktop/r scripts/root biomass - 11082020/ia_mxg2018_root biomass.csv")

nrow(ia) #310 datapoints
ia <- na.omit(ia) 
nrow(ia) #290- these removed points are mostly due to unlabeled samples which must be discarded and not used in analyses. 
par(mfrow=c(2,2))
plot(as.numeric(ia$`theoretical fine roots (g/m^2)`), ylab = expression("fine root biomasss "~g~m^2))
boxplot(as.numeric(ia$`theoretical fine roots (g/m^2)`), ylab = expression("fine root biomasss "~g~m^2))
hist(as.numeric(ia$`theoretical fine roots (g/m^2)`), xlab = expression("fine root biomasss "~g~m^2), main = "histogram")
```

-Change nrate from lbs/acre/yr to Kg/ha/yr
-Change depth key to something informative
```{r}
#switch fertilization labels from lbs/acre to kg/ha
ia$nrate <- ifelse(ia$`fertilization_.level (lbs/acre)` ==0, "0", 
                           ifelse(ia$`fertilization_.level (lbs/acre)` ==100, "112",
                                  ifelse(ia$`fertilization_.level (lbs/acre)` ==200, "224",
                                         ifelse(ia$`fertilization_.level (lbs/acre)` ==300, "336", 
                                                ifelse(ia$`fertilization_.level (lbs/acre)` ==400, "448", "WHAT")))))
#make depth increments easier to comprehend
ia$depth.key <- ifelse(ia$depth.key ==10, "0-10", 
                             ifelse(ia$depth.key ==20, "10-20",
                                    ifelse(ia$depth.key ==30, "20-30",
                                           ifelse(ia$depth.key ==50, "30-50", 
                                                  ifelse(ia$depth.key ==100, "50-100", "WHAT")))))
```

-Subset data for statistical anayses 
    MUCH easier to make new df w/ desired columns than delete unnecessary columns + you can add a column back if necessary
-Rename columns for convenience 
```{r}
ia2 <- (ia[, c("planting.year", "block","depth.key", "theoretical fine roots (g/m^2)", "nrate")])
names(ia2)[names(ia2) == "planting.year"] <- "age" 
names(ia2)[names(ia2) == "depth.key"] <- "depth" 
names(ia2)[names(ia2) == "theoretical fine roots (g/m^2)"] <- "rb" 
```

-Define model variables
-Explore relationship between dependent and independent variables
```{r}
#ia <- subset(ia, select= -c(ia$planting_year == "2015" & ia$depth.key == "50-100"))
age <- as.factor(ia2$age)
nrate <- as.factor(ia2$nrate)
depth <- as.factor(ia2$depth)
rb <- as.numeric(ia2$rb)
block <- as.numeric(ia2$block)
par(mfrow=c(2,2))
boxplot(rb~nrate, ylab=expression("fine root biomass "~(g~m^2)), xlab=expression("fertilization level "~(kg~ha^-1)))
boxplot(rb~age, ylab=expression("fine root biomass "~(g~m^2)), xlab="planting year")
boxplot(rb~block, ylab=expression("fine root biomass "~(g~m^2)), xlab="block")
boxplot(rb~depth, lab=expression("fine root biomass "~(g~m^2)), xlab="depth")
```

-Explore distribution of logged vs. unlogged data
```{r}
log.rb <- log(rb)
par(mfrow=c(2,2))
plot(density.default(x = rb)) #left-skewed 
plot(density.default(x=log.rb)) #good 
ia2$log.rb <- log(as.numeric(ia2$rb))
```

**Part 2: Statistical Analyses**

-Define model and explore residuals
```{r}
ia.mod <- lmer(ia2$log.rb ~ as.factor(ia2$depth) * as.factor(ia2$age) * as.factor(ia2$nrate) + (1|ia2$block)) 
par(mfrow=c(2,2))
plot(ia.mod)
qqnorm(resid(ia.mod)) 
qqline(resid(ia.mod)) 
```

```{r}
z <- car :: Anova(ia.mod)
print(z)
```

-Anova summary into CSV
-Pairwise comparisons 
***Don't run unless you want to 
  1) Rewrite .csv, which may or may not have edits
  2) Wait for pairwise comps. to run which takes forever and output is useless (run chunk by chunk!)
  
```{r, eval=F, echo=T}}
#write.csv(z, file = "E://results//ia_mxg_finebm_anova.csv")
emmeans(ia.mod, pairwise ~ age) 
emmeans(ia.mod, pairwise ~ depth)
emmeans(ia.mod, pairwise ~ nrate) 
emmeans(ia.mod, pairwise ~ nrate | age) 
emmeans(ia.mod, pairwise ~ age | nrate)
emmeans(ia.mod, pairwise ~ depth | age) 
emmeans(ia.mod, pairwise ~ age | depth)
```

**Part 3: Data Summary Statistics and Visualization**

*IND VARIABLE = ESTABLISHMENT YEAR*
-Aggregate data by est. year
-Plot
```{r}
ia.a <- ia2 %>% group_by(age) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.a, aes(x=age, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_continuous("Planting year")+scale_y_continuous(expression("Fine root biomass  "(g/m^2)), breaks=seq(0,500,20), expand = c(0,0))
```

*IND VARIABLE = FERTILIZATION LEVEL*
-Aggregate data by fertiization level
-Plot
```{r}
ia.n <- ia2 %>% group_by(nrate) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.n, aes(x=nrate, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_discrete(expression("Fertilization level "~(kg~ha^-1~yr)))+scale_y_continuous(expression("Fine root biomass  "(g/m^2)), breaks=seq(0,500,20), expand = c(0,0))
```

*IND VARIABLE = SOIL DEPTH INCREMENT*
-Aggregate data by soil depth increment 
-Plot 
```{r}
ia.n <- ia2 %>% group_by(depth) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.n, aes(x=depth, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_discrete("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass  "(g/m^2)), breaks=seq(0,500,20), expand = c(0,0))
```

*IND VARIABLE = SOIL DEPTH INCREMENT X ESTABLISHMENT YEAR*
-Aggregate data by soil depth x planting year
-Plot 
```{r}
ia.da <- ia2 %>% group_by(depth, age) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.da, aes(x=as.factor(depth), y=mean, fill=(as.factor(age))))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4, position = pd)+
    xlab("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,20), expand = c(0,0))+theme(legend.justification = "top")+theme(legend.background = element_rect(linetype = 1, size = 0.5, colour = 1))+scale_fill_manual(values = c("grey10","grey33", "grey91"),"Establishment year")
```

*IND VARIABLE = FERTILIZATION LEVEL X ESTABLISHMENT YEAR*
-Aggregate data by fertilization level x ESTABLISHMENT year
-Plot (two plots for this particular ind. variable combo, useful to see both ways!)
```{r}
ia.na <- ia2 %>% group_by(nrate, age) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.na, aes(x=as.factor(nrate), y=mean, fill=(as.factor(age))))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4, position = pd)+
    xlab(expression("Fertilization level "~(kg~ha^-1~yr)))+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,20), expand = c(0,0))+theme(legend.justification = "top")+theme(legend.background = element_rect(linetype = 1, size = 0.5, colour = 1))+scale_fill_manual(values = c("gray10","grey33", "grey91"),"Establishment year")

ggplot(ia.na, aes(x=as.factor(age), y=mean, fill=(as.factor(nrate))))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4, position = pd)+
    xlab("Establishment year")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,20), expand = c(0,0))+theme(legend.justification = "top")+theme(legend.background = element_rect(linetype = 1, size = 0.5, colour = 1))+scale_fill_manual(values = c("gray10","grey33", "grey91", "grey69", "grey52"), expression("Fertilization level "~(kg~ha^-1~yr)))
```

-Save summary stats as .csvs 
***Don't run this unless you've made changes to .csvs that you WANT to save
```{r, eval=F, echo=T}
#write.csv(ia.all, file = "E://results//ia_mxg_bmF_all_means.csv")
#write.csv(ia.a, file = "E://results//ia_mxg_bmF_age_means.csv")
#write.csv(ia.n, file = "E://results//ia_mxg_bmF_nrate_means.csv")
#write.csb(ia.d, file= "E://results//ia_mxg_bmF_depth_means.csv")
```


