 ---
title: "Fine root biomass- Miscanthus x giganteus - IA"
author: "Ryan Quinn"
date: "March 7, 2020"
output: html_document
---

**Part 1: Data Cleaning and Early Exploration**

Notes: 
  -The dataset for this 1m analysis is much smaller than the data for all the depth increments seperately. To allow the data to be stored on the same .csv, all cells in the data column not used to store data were entered with the word "Delete", to allow for easy condensing of the spreadsheet for analyses.  
  -One outlier exists in this dataset. Removing vs. not removing does not change the statistical results of this data. I do not like to remove the outliers unless they're 

-Load libraries
-Define functions needed throughout script
-Load data, remove missing datapoints 
-Early data exploration
```{r}
library(readr)
library(lmerTest)
library(emmeans)
library(ggplot2)
library(dplyr) #summary statistics 
library(viridis) 
se.error <- function(x)sd(x)/sqrt(length(x)) #calculate SE during data aggregation 
pd <- position_dodge(width = 1) #position error bars on  figures

ia <- read_csv("C:/Users/ryan_/OneDrive/Desktop/r scripts/root biomass - 11082020/ia_mxg2018_root biomass.csv")

ia$`theoretical fine roots in 1m (g/m2)` <- ifelse(ia$`theoretical fine roots in 1m (g/m2)` == "Delete", NA, ia$`theoretical fine roots in 1m (g/m2)`) #VERY IMPORTANt 
nrow(ia) 
ia <- na.omit(ia) 
nrow(ia)
par(mfrow=c(2,2))
plot(as.numeric(ia$`theoretical fine roots in 1m (g/m2)`))
boxplot(as.numeric(ia$`theoretical fine roots in 1m (g/m2)`)) #one lone outlier. Don't remove. 
hist(as.numeric(ia$`theoretical fine roots in 1m (g/m2)`), main="FRB in 1m- histogram", xlab=expression("FRB in 1m "~(g~m^2)))
```

-Change nrate from lbs/acre/yr to Kg/ha/yr
```{r}
#switch fertilization labels from lbs/acre to kg/ha
ia$nrate <- ifelse(ia$`fertilization_.level (lbs/acre)` ==0, "0", 
                           ifelse(ia$`fertilization_.level (lbs/acre)` ==100, "112",
                                  ifelse(ia$`fertilization_.level (lbs/acre)` ==200, "224",
                                         ifelse(ia$`fertilization_.level (lbs/acre)` ==300, "336", 
                                                ifelse(ia$`fertilization_.level (lbs/acre)` ==400, "448", "WHAT")))))
```

-Subset data for statistical anayses 
    MUCH easier to make new df w/ desired columns than delete unnecessary columns + you can add a column back if necessary
-Rename columns for convenience 
```{r}
ia2 <- (ia[, c("planting.year", "block", "theoretical fine roots in 1m (g/m2)", "nrate")])
names(ia2)[names(ia2) == "planting.year"] <- "age" 
names(ia2)[names(ia2) == "theoretical fine roots in 1m (g/m2)"] <- "rb" 
```

-Define model variables
-Explore relationship between dependent and independent variables
```{r}
#ia <- subset(ia, select= -c(ia$planting_year == "2015" & ia$depth.key == "50-100"))
age <- as.factor(ia2$age)
nrate <- as.factor(ia2$nrate)
rb <- as.numeric(ia2$rb)
block <- as.numeric(ia2$block)
par(mfrow=c(2,2))
boxplot(rb~nrate)
boxplot(rb~age)
boxplot(rb~age)
boxplot(rb~block)
```

-Explore distribution of logged vs. unlogged data
```{r}
log.rb <- log(rb)
par(mfrow=c(2,2))
plot(density.default(x = rb)) #left-skewed 
plot(density.default(x=log.rb)) #good 
```

**Part 2: Statistical Analyses**

-Define model and explore residuals
```{r}
ia.mod <- lmer(log.rb ~ nrate * age * (1|block)) 
par(mfrow=c(2,2))
plot(ia.mod)
qqnorm(resid(ia.mod)) 
qqline(resid(ia.mod)) 
```

```{r}
z <- car :: Anova(ia.mod)
print(z)
```

-Anova summary into CSV
-Pairwise comparisons 
***Don't run unless you want to 
  1) Rewrite .csv, which may or may not have edits
  2) Wait for pairwise comps. to run which takes forever and output is useless (run chunk by chunk!)
  reval=F, echo=T}
```{r, eval=F, echo-T}
#write.csv(z, file = "E://results//ia_mxg_finebm_anova.csv")
emmeans(ia.mod, pairwise ~ age) 
emmeans(ia.mod, pairwise ~ nrate) 
emmeans(ia.mod, pairwise ~ nrate | age) 
emmeans(ia.mod, pairwise ~ age | nrate)
```

**Part 3: Data Summary Statistics and Visualization**


**IND VARIABLE = ESTABLISHMENT YEAR**
-Aggregate data by est. year
-Plot
```{r}
ia.a <- ia2 %>% group_by(age) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.a, aes(x=age, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 16)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4)+
  scale_x_continuous("Establishment year")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,1000,100), expand = c(0,0))
```


**IND VARIABLE = FERTILIZATION LEVEL**
-Aggregate data by fertiization level
-Plot
```{r}
ia.n <- ia2 %>% group_by(nrate) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.n, aes(x=nrate, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 16)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4)+
  scale_x_discrete(expression("N fertilization" ~(kg~N~ha^-1~yr^-1)))+scale_y_continuous(expression("Fine root biomass "(g~m^2)), breaks=seq(0,700,100), limits = c(0,700), expand = c(0,0))
```

**IND VARIABLE = ESTABLISHMENT YEAR X FERTILIZATION LEVEL**
-Aggregate data by genotype
-Plot fine root biomass by genotype
-Plot: AGE & nrate
```{r}
ia.an <- ia2 %>% group_by(nrate, age) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(ia.an, aes(age, mean, fill=(nrate)))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
      #scale_fill_discrete(name="depth increment")+ 
          scale_fill_viridis(expression("N fertilization "(kg~N~ha^-1~yr^1)), discrete = T, direction = -1, option = "viridis")+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.3, position = pd)+
    xlab("Establishment year")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,1000,100), expand = c(0,0))+theme(legend.justification = "top")

ggplot(ia.an, aes(nrate, mean, fill=(as.factor(age))))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
      #scale_fill_discrete(name="depth increment")+ 
          scale_fill_viridis("Establishment year", discrete = T, direction = -1, option = "viridis")+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.3, position = pd)+
    scale_x_discrete(expression("N fertilization "(kg~N~ha^-1~yr^1)))+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,1000,100), expand = c(0,0))+theme(legend.justification = "top")

write.csv(ia.an, file = "E://results//mxg_nrate_age_1m.csv")
```

**PART 4: Wrangle data for comparison of mean root resp (all fertilization levels grouped, and just controls ( 0 kg/ha))**
**eventually move this into another spreadsheet, call this spreadsheet before combining the data. 

-Calc. g mean +/- SE frb in all Miscanthus plots over all fertilization levels and sampling dates
```{r}
frb.mxg <- NA
frb.mxg <- as.data.frame(frb.mxg) 
frb.mxg$mean <- mean(as.numeric(ia2$rb))
frb.mxg$n <- length(ia2$rb)
frb.mxg$se <- se.error(ia2$rb)
frb.mxg$crop <- "Miscanthus"
frb.mxg
```

-Calc. mean +/- SE of 0 kg/ha fertilization level for g fine root biomass in Miscanthus across all sampling dates 
```{r}
frb.0.mxg <- subset(ia2, ia2$nrate == 0)
frb.0.mxg <- as.numeric(frb.0.mxg$rb) #get rid of everything else in df, only need frb 

frb.mxg.0 <- NA
frb.mxg.0 <- as.data.frame(frb.mxg.0) 
frb.mxg.0$mean <-  mean(frb.0.mxg)
frb.mxg.0$n <- length(frb.0.mxg)
frb.mxg.0$se <- se.error(frb.0.mxg)
frb.mxg.0$crop <- "Miscanthus"
frb.mxg.0
```
