---
title: "Fine root biomass - sorghum -  IL"
author: "Ryan Quinn"
date: "March 22, 2020"
output: html_document
---

**Part 1: Data Cleaning and Early Exploration**

**Notes**: 
  -Earlier data exploration examined the effects of outlier removal on the model. Removing outliers removes three datapoints all from the 0 kg/ha fertilization level (nrate). Through expert elication we decided the datapoints were "real" data and represent logical amounts of fine RB expected in the 0 fertilization level.
  -For the purposes of streamlining publication, only data with statistical significance is visualized 
  -This data IS NOT PUBLICALLY AVAILABLE and will not be until the publication is in press. However, these scripts may help someone currently processing soil C flux data. 

-Load libraries
-Define functions needed throughout script
-Load data, remove missing datapoints
-Basic exploration of data distribution, shape, etc. 
```{r}
library(readr)
library(lmerTest)
library(emmeans)
library(ggplot2) 
#library(viridis) #went with colorcoding by x variable instead of colorblind friendly for presentation b/c nobody on committee colorblind
library(plyr) #must unload this before using dplyr
library(dplyr) #summary statistics 


se.error <- function(x)sd(x)/sqrt(length(x)) #calculate SE during data aggregation 
sumstat <- function(x) c(mean(x), se.error(x), length((x)))
pd <- position_dodge(width = 1) #position error bars on  figures

il <- read_csv("C:/Users/ryan_/OneDrive/Desktop/r scripts/root biomass - 11082020/il_sorghum2018_root biomass.csv") 

nrow(il) #240 rows before removing missing datapoints
il <- na.omit(il)
nrow(il) #200 rows once the missing values are removed from the dataset (the "missing data" is not actually missing but was never collected purposefully for a variety of reasons specific to the experiment, and/or the data collected was not accurately labels (no plot #, etc.)).
par(mfrow=c(2,2))
plot(as.numeric(il$`theoretical fine roots (g/m^2)`))
boxplot(as.numeric(il$`theoretical fine roots (g/m^2)`)) #3 outliers 
hist(as.numeric(il$`theoretical fine roots (g/m^2)`))
```

-Change nrate from lbs/acre/yr to Kg/ha/yr
```{r}
il$nrate <- ifelse(il$`fertilization level (lbs/acre)` ==0, "0", 
                           ifelse(il$`fertilization level (lbs/acre)` ==150, "168", "WHAT"))
```

-Subset data for statistical anayses 
    MUCH easier to make new df w/ desired columns than delete unnecessary columns + you can add a column back if necessary
-Rename columns for convenience 
```{r}
il2 <- (il[, c("interrow (IR)/intrarow (OR)", "genotype", "nrate","theoretical depth (cm)", "theoretical fine roots (g/m^2)")])
#il2 <- as.data.frame(rename(il2, c("interrow (IR)/intrarow (OR)" = "row", "nrate" = "nrate", "theoretical fine roots (g/m^2)" = "rb", "theoretical depth (cm)" = "depth", "genotype" = "genotype")))
names(il2)[names(il2) == "interrow (IR)/intrarow (OR)"] <- "row" 
names(il2)[names(il2) == "theoretical depth (cm)"] <- "depth"  
names(il2)[names(il2) == "theoretical fine roots (g/m^2)"] <- "rb"  

```


-Define model variables
-Explore relationship between dependent and independent variables
```{r}
depth <- as.factor(il2$depth)
genotype <- as.factor(il2$genotype)
nrate <- as.factor(il2$nrate)
row <- as.factor(il2$row)
rb <- as.numeric(il2$rb)
par(mfrow=c(2,2))
boxplot(rb~depth) #fix this x axis

boxplot(rb~nrate) 
boxplot(rb~genotype) 
boxplot(rb~row)
```

-Explore distribution of logged vs. unlogged data
```{r}
log.rb <- log(rb)
par(mfrow=c(1,2))
plot(density.default(x = rb)) #left-skewed 
plot(density.default(x=log.rb))  
```

**Part 2: Statistical Analyses**


-Define model and explore residuals
```{r}
il.mod <- lm(rb ~ nrate * row * genotype * depth) 
qqnorm(resid(il.mod)) 
qqline(resid(il.mod)) 
```

```{r}
z <- car :: Anova(il.mod)
print(z)
```

-Anova summary into CSV
-Pairwise comparisons 
***Don't run unless you want to 
  1) Rewrite .csv, which may or may not have edits
  2) Wait for pairwise comps. to run which takes forever and output is useless (run chunk by chunk!)
  
```{r, eval=F, echo=T}
#write.csv(z, file = "E://results//ia_mxg_finebm_anova.csv")
emmeans(il.mod, pairwise ~ depth) 
emmeans(il.mod, pairwise ~ genotype) 
emmeans(il.mod, pairwise ~ row) 
emmeans(il.mod, pairwise ~ depth | row) 
emmeans(il.mod, pairwise ~ row | depth)
```

**Part 3: Data Summary Statistics and Visualization**
                                                
```{r}
detach(package:plyr) #if you don't run this, the dplyr summarization tables don't group
```                              
#row.m <- as.data.frame(do.call( rbind, tapply(rb, list(row), sumstat))) #alternative method to summarize data 

*IND VARIABLE = ROW LOCATION*
-Aggregate data by row
-Rename x-axis labels for plot
-Plot fine root biomass by row location
```{r}
df.r <- il2 %>% group_by(row) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

df.r$row <- ifelse(df.r$row == "IR", "Inter-row",
                   ifelse(df.r$row == "OR", "Intra-row", "WHAT"))

ggplot(df.r, aes(x=row, y=mean))+
  geom_col(fill = "seagreen4")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_discrete("Row location")+scale_y_continuous(expression("Fine root biomass "(g~m^2)), breaks=seq(0,40,10), limits = c(0,40), expand = c(0,0))
```

*IND VARIABLE = DEPTH*
-Aggregate data by depth
-Plot fine root biomass by depth
```{r}
df.d <- il2 %>% group_by(depth) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))
ggplot(df.d, aes(x=depth, y=mean))+
  geom_col(fill = "seagreen4")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_discrete("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass "(g~m^2)), breaks=seq(0,80,10), limits = c(0,80), expand = c(0,0))
```

*IND VARIABLE = GENOTYPE*
-Aggregate data by genotype
-Plot fine root biomass by genotype
```{r}
df.g <- il2 %>% group_by(genotype) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))
ggplot(df.g, aes(x=genotype, y=mean))+
  geom_col(fill = "seagreen4")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.2)+
  scale_x_discrete("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass "(g~m^2)), breaks=seq(0,40,10), limits = c(0,40), expand = c(0,0))
```

*IND VARIABLE = Row x Depth*
-Aggregate data by genotype
-Plot fine root biomass by genotype
```{r}
df.dr <- il2 %>% group_by(row, depth) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(df.dr, aes(x=depth, y=mean, fill=(row)))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4, position = pd)+
    xlab("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,20), expand = c(0,0))+theme(legend.justification = "top")+theme(legend.background = element_rect(linetype = 1, size = 0.5, colour = 1))+scale_fill_manual(values = c("seagreen1","seagreen4"),"Row location")

write.csv(df.dr, file = "E://results//frb_row_depth.csv")

```

*IND VARIABLE = Nrate x Depth*
-Aggregate data by genotype
-Plot fine root biomass by genotype
```{r}
df.dn <- il2 %>% group_by(nrate, depth) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(df.dn, aes(x=depth, y=mean, fill=(as.character(nrate))))+
geom_col(position="dodge")+theme_classic(base_size = 16)+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4, position = pd)+
    xlab("Soil depth (cm)")+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,20), expand = c(0,0))+theme(legend.justification = "top")+theme(legend.background = element_rect(linetype = 1, size = 0.5, colour = 1))+scale_fill_manual(values = c("seagreen1","seagreen4"),expression("N fertilization level "(~kg~ha^-1)))
```
