---
title: "RB-Fine (IL srg)"
author: "Ryan Quinn"
date: "March 22, 2020"
output: html_document
---

**Part 1: Data Cleaning and Early Exploration**

Notes: 
  -Earlier data exploration examined the effects of outlier removal on the model. Removing outliers removes most 0-10cm depth bm datapoints So, we do not remove outliers for this analysis. 

-Load libraries
-Define functions needed throughout script
-Load data, remove missing datapoints (Empty cells with "Delete" are not missing data, but a code to delete cells which must remain filled for other analyses where we're looking at depth distribution and not total FRB in 1m)
```{r}
library(readr)
library(lmerTest)
library(emmeans)
library(ggplot2)
library(dplyr) #summary statistics 
library(viridis)
se.error <- function(x)sd(x)/sqrt(length(x)) #calculate SE during data aggregation 
pd <- position_dodge(width = 1) #position error bars on  figures
il <- read_csv("C:/Users/ryan_/OneDrive/Desktop/r scripts/root biomass - 11082020/il_sorghum2018_root biomass.csv")
il$`theoretical fine roots in 1m (g/m^2)` <- ifelse(il$`theoretical fine roots in 1m (g/m^2)` == "Delete", NA, il$`theoretical fine roots in 1m (g/m^2)`)
```

-Early data exploration 
```{r}
nrow(il) 
il <- na.omit(il)
nrow(il)
par(mfrow=c(2,2))
plot(as.numeric(il$`theoretical fine roots in 1m (g/m^2)`))
boxplot(as.numeric(il$`fine roots in sample (g/m^2)`))
hist(as.numeric(il$`fine roots in sample (g/m^2)`))
```

-Change nrate from lbs/acre/yr to Kg/ha/yr
```{r}
#switch fertilization labels from lbs/acre to kg/ha
il$nrate <- ifelse(il$`fertilization level (lbs/acre)` ==0, "0", 
                           ifelse(il$`fertilization level (lbs/acre)` ==150, "168", "WHAT"))
```

-Subset data for statistical anayses 
    MUCH easier to make new df w/ desired columns than delete unnecessary columns + you can add a column back if necessary
-Rename columns for convenience 
```{r}
il2 <- (il[, c("theoretical fine roots in 1m (g/m^2)", "nrate", "interrow (IR)/intrarow (OR)", "genotype")])
names(il2)[names(il2) == "interrow (IR)/intrarow (OR)"] <- "row" 
names(il2)[names(il2) == "theoretical fine roots in 1m (g/m^2)"] <- "rb" 
```

-Define model variables
-Explore relationship between dependent and independent variables
```{r}
nrate <- as.factor(il$nrate)
rb <- as.numeric(il$`theoretical fine roots in 1m (g/m^2)`)
genotype <- as.factor(il$genotype)
row <- as.factor(il$`interrow (IR)/intrarow (OR)`)
par(mfrow=c(2,2))
boxplot(rb~nrate)
boxplot(rb~genotype)
boxplot(rb~row)
```

-Explore distribution of logged vs. unlogged data
```{r}
log.rb <- log(rb)
par(mfrow=c(2,2))
plot(density.default(x = rb)) #left-skewed 
plot(density.default(x=log.rb)) #good 
```

**Part 2: Statistical Analyses**

-Define model and explore residuals
```{r}
il.mod <- lm(log.rb ~ nrate * row * genotype) 
par(mfrow=c(2,2))
qqnorm(resid(il.mod)) 
qqline(resid(il.mod)) 
```

```{r}
z <- car :: Anova(il.mod)
print(z)
```

-Anova summary into CSV
-Pairwise comparisons 
***Don't run unless you want to 
  1) Rewrite .csv, which may or may not have edits
  2) Wait for pairwise comps. to run which takes forever and output is useless (run chunk by chunk!)
  reval=F, echo=T}
```{r}
#write.csv(z, file = "results/mxg_frb_anova.csv") #change this pathway when data is publically availabel to write into 
emmeans(il.mod, pairwise ~ genotype) 
emmeans(il.mod, pairwise ~ depth | year)  
emmeans(il.mod, pairwise ~ row) 
emmeans(il.mod, pairwise ~ depth | row) 
emmeans(il.mod, pairwise ~ row | depth)
```

**Part 3: Data Summary Statistics and Visualization**

**IND VARIABLE = row location**
-Aggregate data by location to plant within/outside plant row
-Plot
```{r}
df.r <- il2 %>% group_by(row) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

df.r$row <- ifelse(df.r$row == "IR", "Inter",
                   ifelse(df.r$row == "OR", "Intra", "WHAT"))

ggplot(df.r, aes(x=row, y=mean))+
  geom_col(position = "dodge")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4)+
  scale_x_discrete("Proximity to plant row")+scale_y_continuous(expression("Fine root biomass "(g~m^-2)), breaks=seq(0,500,40), expand = c(0,0))#+scale_fill_manual(values = c("seagreen4","seagreen4"))
```

**IND VARIABLE = GENOTYPE**
-Aggregate data by genotype
-Plot
```{r}
df.g <- il2 %>% group_by(genotype) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))

ggplot(df.g, aes(x=genotype, y=mean))+
  geom_col(position = "dodge", fill="seagreen4")+theme_classic(base_size = 16)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4)+
  scale_x_discrete("Genotype")+scale_y_continuous(expression("Fine root biomass "(g~m^-2)), breaks=seq(0,500,40), expand = c(0,0))
#write.csv(df.g, file = "E://results//sorg_frb_geno_1m.csv") 
```

**IND VARIABLE = FERTILIZATION LEVEL**
-Aggregate data by fertiization level
-Plot
```{r}
df.n <- il2 %>% group_by(nrate) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))
ggplot(df.n, aes(x=nrate, y=mean))+
  geom_col(position = "dodge", fill="seagreen4")+theme_classic(base_size = 13)+
  geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.4)+
  scale_x_discrete(expression("N fertilization" ~(kg~ha^-1~yr^-1)))+scale_y_continuous(expression("Fine root biomass "~(g~m^-2)), breaks=seq(0,500,30), expand = c(0,0))
```

**IND VARIABLE = FERTILIZATION LEVEL x ROW **
-Aggregate data by fertiization level x row location
-Plot
```{r}
df.n <- il2 %>% group_by(nrate, row) %>% summarize(mean = mean(as.numeric(rb)), 
                                    se = se.error(rb), 
                                    n = length(rb))
df.r$row <- ifelse(df.r$row == "IR", "Inter",
                   ifelse(df.r$row == "OR", "Intra", "WHAT"))

ggplot(df.n, aes(nrate, mean, fill=(row)))+
geom_col(position="dodge")+theme_classic(base_size = 13)+
          scale_fill_viridis(expression("Row proximity"), discrete = T, direction = -1, option = "viridis")+
geom_errorbar(aes(ymin= mean-se, ymax=mean+se), color = "black", width=.3, position = pd)+
    scale_x_discrete(expression("N fertilization "(kg~ha^-1~yr^1)))+
                       scale_y_continuous(expression("fine root biomass "~(g~m^-2)), breaks=seq(0,1000,100), expand = c(0,0))+theme(legend.justification = "top")
```

**PART 4: Wrangle data for comparison of mean root resp (all fertilization levels grouped, and just controls ( 0 kg/ha))**
**eventually move this into another spreadsheet, call this spreadsheet before combining the data. 

-Calc. g mean +/- SE frb in all Sorghum plots over all fertilization levels and sampling dates
```{r}
srg.frb <- NA
srg.frb <- as.data.frame(srg.frb) 
srg.frb$mean <- mean(as.numeric(il2$rb))
srg.frb$n <- length(il2$rb)
srg.frb$se <- se.error(il2$rb)
srg.frb$crop <- "Sorghum"
srg.frb
```

-Calc. mean +/- SE of 0 kg/ha fertilization level for g fine root biomass in Sorghum across all sampling dates 
```{r}
frb.0.srg <- subset(il2, il2$nrate == 0)
frb.0.srg <- as.numeric(frb.0.srg$rb) #get rid of everything else in df, only need frb 

frb.srg.0 <- NA
frb.srg.0 <- as.data.frame(frb.srg.0) 
frb.srg.0$mean <-  mean(frb.0.srg)
frb.srg.0$n <- length(frb.0.srg)
frb.srg.0$se <- se.error(frb.0.srg)
frb.srg.0$crop <- "Sorghum"
frb.srg.0
```